The calibration of the Heston stochastic volatility model constitutes a non-trivial inverse problem, typically constrained by the computational intensity of standard numerical pricing methods. This study evaluates the longitudinal robustness and accuracy of a \ac{ddn} employed as a surrogate pricing engine within a hybrid calibration framework. The \ac{ddn} is trained exclusively on a synthetic dataset generated via \ac{lhs} using Sobolev training, ensuring no prior exposure to historical market data. The network approximates both option prices and their partial derivatives, thereby facilitating rapid convergence within a classical L-BFGS-B optimization routine. 

To validate the methodology, a comprehensive empirical backtest was performed on \ac{aapl} options from 2016 to 2023. For each trading day, the option chain was partitioned into a calibration set (80\%) for parameter inference and a distinct hold-out validation set (20\%) for pricing assessment. The framework demonstrates high fidelity, achieving an average \ac{mre} of 4.34\% on the hold-out validation set. This metric closely tracks the calibration set \ac{mre} of 4.14\%, indicating that the inferred parameters generalize well to unseen contracts despite the \ac{ddn} relying solely on synthetic pre-training. While calibration error exhibits a positive correlation with realized market volatility, notably during the COVID-19 market crash, the model maintains stability and demonstrates rapid error mean reversion. Furthermore, the methodology proves resilient across varying macroeconomic conditions, including regimes characterized by negative market-implied risk-free rates. These findings confirm that \ac{ddn}-based calibration effectively resolves the trade-off between computational efficiency and pricing accuracy, offering a scalable solution for real-time risk management.

\textit{Keywords:} Heston Model, Deep Differential Networks, Sobolev Training, Stochastic Volatility, Model Calibration, Deep Learning, Quantitative Finance

\textit{JEL Classification:} G12, C45